The \dealii Library, Version 9.6

### 2024-09-18 9.6/paper.tex
Resolve one TODO.

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
  \item deal.II now requires and uses of C++17.
  \item We have continued to make progress in supporting simplex and
    mixed meshes -- mesh types that \dealii{} has traditionally not
    supported at all. Specifically, the current release uses
    better strategies for refinement of tetrahedra that result in
    better-shaped child cells. It also contains
    support for cubic finite elements on simplices.
  \item The \texttt{FE\_NedelecSZ} class that contains our
    implementation of the N\'ed\'elec element using the orientation
    scheme of \cite{Zag06} now supports the computation of hanging
    node constraints for locally refined meshes.
    \todo[inline]{Sebastian: Is there anything more to say here, maybe some reference? If
      you want, we can also list this point in the list above, and add
      a whole section 2.x for it.}
  \item changes of interface of AffineConstraints
    \todo[inline]{Wolfgang to write}
  \item \dealii{} internally needs to keep track of the relative
    orientations of the coordinate systems associated with neighboring
    cells, as well as of the coordinate systems of faces and edges
    within the coordinate system of the cells they are part
    of. Historically, different places within the library that had to
    do that had grown their own, independent ways of representating
    this information, much of which was difficult to understand.
    We have now cleaned up this area. Specifically, instead of using three
    Booleans to define relative orientations for hexedra (\textit{orientation},
    \textit{flip}, and \textit{rotation}), we have internally merged these three values into a single variable of type
     \texttt{char} -- i.e., the smallest integer data type that has at
     least three bits to represent the information. Because the way to represent
     orientations is part of the public interface, the change causes
     some incompatibilities in types and function signatures.
  \todo[inline]{DavidW: Please describe some details here. -- When
    done, please remove this comment.}
\item The \texttt{SolverGMRES} class now offers a third orthogonalization
  method, the classical Gram--Schmidt method with delayed orthogonalization
  \cite{Bielich2022}. Furthermore, the solver now specifies the maximal basis
  size of the Arnoldi basis, rather than the number of auxiliary
  vectors. Some changes have also been made to the GMRES and F-GMRES
  implementations, making them use the same underlying kernels as much as possible.
\end{itemize}
%
The changelog lists more than 180 other features and bugfixes.

>>>>

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
  \item deal.II now requires and uses of C++17.
  \item We have continued to make progress in supporting simplex and
    mixed meshes -- mesh types that \dealii{} has traditionally not
    supported at all. Specifically, the current release uses
    better strategies for refinement of tetrahedra that result in
    better-shaped child cells. It also contains
    support for cubic finite elements on simplices.
  \item The \texttt{FE\_NedelecSZ} class that contains our
    implementation of the N\'ed\'elec element using the orientation
    scheme of \cite{Zag06} now supports the computation of hanging
    node constraints for locally refined meshes.
    \todo[inline]{Sebastian: Is there anything more to say here, maybe some reference? If
      you want, we can also list this point in the list above, and add
      a whole section 2.x for it.}
  \item The \texttt{AffineConstraints} class stores and processes
    constraints on degrees of freedom in \dealii{}. Such constraints
    can be of the form $x_3 = \frac 12 x_{14} + \frac 12 x_{15}$ as is
    common when using hanging node constraints (here, $x_3$ could
    represent the value of the solution at a hanging node, and $x_{14}$
    and $x_{15}$ are the values of the two adjacent degrees of freedom on
    the parent edge that contains the hanging node); of they can be of
    the form $x_{12}=42$ as is common when using Dirichlet boundary
    conditions.

    Since very early in the history of the library, the
    \texttt{AffineConstraints} class interface required building such
    constraints in multiple steps: First, one declared a degree of
    freedom as constrained; then one added the dependencies one after
    the other (e.g., by adding pairs $(\frac 12, 13)$ and $(\frac 12,
    14)$ in the hanging node example above); then one added
    inhomogeneities (by setting it to $42$ in the Dirichlet example
    above). This piecemeal approach is cumbersome and prevents the
    library from performing certain error checking steps because a
    constraint is not known to be completely built at any given point. The new
    \textt{add\_constraint()} function now allows defining a
    constraint in one step.
  \item \dealii{} internally needs to keep track of the relative
    orientations of the coordinate systems associated with neighboring
    cells, as well as of the coordinate systems of faces and edges
    within the coordinate system of the cells they are part
    of. Historically, different places within the library that had to
    do that had grown their own, independent ways of representating
    this information, much of which was difficult to understand.
    We have now cleaned up this area. Specifically, instead of using three
    Booleans to define relative orientations for hexedra (\textit{orientation},
    \textit{flip}, and \textit{rotation}), we have internally merged these three values into a single variable of type
     \texttt{char} -- i.e., the smallest integer data type that has at
     least three bits to represent the information. Because the way to represent
     orientations is part of the public interface, the change causes
     some incompatibilities in types and function signatures.
  \todo[inline]{DavidW: Please describe some details here. -- When
    done, please remove this comment.}
\item The \texttt{SolverGMRES} class now offers a third orthogonalization
  method, the classical Gram--Schmidt method with delayed orthogonalization
  \cite{Bielich2022}. Furthermore, the solver now specifies the maximal basis
  size of the Arnoldi basis, rather than the number of auxiliary
  vectors. Some changes have also been made to the GMRES and F-GMRES
  implementations, making them use the same underlying kernels as much as possible.
\end{itemize}
%
The changelog lists more than 180 other features and bugfixes.
